---
title: "Association of obstructive sleep apnea with brain volummetry and cognition in de novo Parkinson's disease"
author: "Josef Mana"
date: "`r Sys.Date()`"
format:
  pdf:
    df-print: kable
bibliography: references.bib
---

```{r}
#| label: envir
#| warning: false
#| echo: false

library(tidyverse)
library(here)
library(gt)


```


# Methods

## Neuropsychological assessment

All patients (PD) and healthy controls (HC) were administered a battery of neuropsychological tests at enrollment to the study and a subsample of participants were administered the same tests at re-test four years after the enrollment. The battery included assessment of (i) declarative memory via Rey Auditory Verbal Learing Test (RAVLT) [@bezdicek2014; @frydrychova2018]; (ii) attention via Trail Making Test, part A (TMT-A) [@bezdicek2012; @bezdicek2017a], and dot colour naming (PST-D) as well as naming colour of neutral words (PST-W) conditions from Prague Stroop Test [@bezdicek2015a]; (iii) executive function via Trail Making Test, part B [@bezdicek2012; @bezdicek2017a], and Prague Stroop Test, interference condition (i.e., naming colour of contrasting colour words, PST-C) [@bezdicek2015a]; and (iv) processing speed via Grooved Pegboard Test (GPT) [@klove1963]. The patients were further examined using tests from the standard International Parkinson and Movement Disorder Society (MDS) neuropsychological battery at Level II for mild cognitive impairment in Parkinson’s disease (PD-MCI) [@litvan2012;  @bezdicek2017]. The Czech normative calculator established by @bezdicek2017 was used to assign PD-MCI diagnosis to each PD patient separately at enrollment and retest. Finally, all participants were administered Montreal cognitive assessment (MoCA) [@kopecek2017; @nasreddine2005] annually for cognitive screening.

## Statistical analyses

All demographic (age, education, gender) and descriptive clinical (iRBD, MoCA, BMI, BDI-II, MDS-UPDRS III, disease duration) variables were described by their mean and standard deviation if continuous and frequency if nominal separately for HC OSA-, HC OSA+, PD OSA-, and PD OSA+ groups of participants, and compared by Gaussian (continuous variables) or logistic (binary variables) regression with group (PD vs. HC), OSA (OSA+ vs. OSA-) and their interaction as predictors. As we did not aim to control type I error rate in these analyses, decision threshold for claiming statistically significant difference in demographic and decscriptive clinical variables was set at p < .05 without adjustment for multiple comparisons.

### Cortical thickness

*4 Fíla*

### Subcortical volummetry

The strength of association of subcortical structures' volume with group (PD vs HC) and OSA (OSA+ vs OSA-) was evaluated by a set of univariate linear regressions fitted to data via QR decomposition. Each subcortical structure's volume was regressed on group, OSA and their interaction as exposures of interest, and years of age, gender and TIV as additive linear covariates. All continuous variables were standardised (i.e., mean-centred and scaled by their in-sample standard deviation) before entering the analysis, and nominal variables were entered via sum coding. Each regression was tested for deviation from normality of residuals via the Shapiro-Wilk test and deviation from homoscedasticity via the Breusch-Pagan test. Within the classical Neyman-Pearson hypothesis testing framework, we selected a decision threshold for rejecting the hypothesis of zero association between subcortical structure's volume and group/OSA by adjusting for 5% false discovery rate (FDR) [@benjamini1995] in tests of all main effects and interactions of interest (i.e., coefficients related to the group and OSA variables) across all subcortical structures examined in this section. The test statistics consisted of *t*-values for each relevant regression coefficient calculated by lm() function in R software for statistical computing [@rsoft]. The primary estimand of interest, i.e., the interaction between group and OSA, was further characterised by calculating simple main effects of OSA within each group and the difference between these simple main effects via the avg_comparisons() function in the "marginaleffects" R package [@marginaleffects].

Based on the results of the primary analyses outlined above, we further explored association of hippocampal fields with group and OSA in a series of post-hoc univariate regressions with identical specification as described above. The only difference between the primary analyses of the gross subcortical structures described above, and the analysis of hippocampal fields was that in the latter case, we focused on the group/OSA interaction only. The decision threshold for rejecting the hypothesis of zero group/OSA interaction was set according to adjustment for 5% FDR within this set of analyses.

Finally, to reflect current recommendations of the American Statistical Association which advised against basing scientific conclusions on whether a *p*-value passes a specific threshold [@wasserstein2016], we also calculated Shannon information (i.e., s-values) $\text{-}log_2{(p)}$. S-value is cognitive tool to help researchers intuitively evaluate strength of evidence against a null hypothesis contained in the results as equivalent to the number consecutive "heads" tosses that would provide the same amount of evidence against the null hypothesis that the coin is fair [@greenland2019; @cole2021].

### Cognitive variables

The distribution of cognitive performance conditional on group, OSA, and measurement occasion (enrollment vs retest) was evaluated using a set of independent Bayesian linear mixed models (LMMs) with cognitive test scores regressed on group, OSA, occasion, and their interactions as fixed effects on group-level and participant-specific random intercepts. All outcomes were standardised before entering the analysis and response time variables were log-transformed before standardising. Contrasts described by @rouder2012 as implemented in contr.equalprior() function from the "bayestestR" R package [@makowski2019] were specified for all group-level parameters to ensure that all pairwise prior differences are centred around zero. Weakly informative priors, i.e. *Normal(0,1)* for group-level parameters and *Exponential(1)* for participant-level and residual variances, were specified to ensure the model converges to reasonable parameter values. Employing LMMs allowed us to use all data without requiring the participants to have both enrollment and retest measurements.

To ensure the results are informed by data, prior and likelihood sensitivity of posterior estimates was evaluated via the powerscale_sensitivity() function from "priorsense" R package [@kallioinen2023]. Moreover, to evaluate model fit, posterior predictive means and SDs of the model within each combination of group, OSA, and assessment were visually compared to values observed in the data. All LMMs were fitted using Stan’s (version 2.21.0) build-in Hamiltonian Monte Carlo sampler accessed via R version `r with( version, paste(major,minor,sep=".") )` using package “brms” [@bürkner2017; @rsoft; @stan]. Four parallel chains were run each for 2,000 iterations for each LMM with the first 1,000 iterations serving as a warm-up. Convergence was checked numerically by inspection of the $\hat{R}s$ and visually by inspection of trace plots.

After establishing model fit, the results were summarised by computing pairwise comparisons based on main effects, two-way interactions and three-way interactions between group, OSA and occasion. These comparisons were then described by their medians, 95% equal-tailed posterior probability intervals (PPIs), and probability of direction (*pd*, i.e., the certainty associated with the most probable direction of the effect) on the original outcome scale. Since this portion of our analysis was purely exploratory and was not set-up to formally test any hypothesis, we did not set any decision threshold regarding reported posterior comparisons but instead interpreted *pd* as continous index of effect existence probability following guidelines of @makowski2019b whereby *pd* ≥ 95% indicates uncertain effect, *pd* > 95% indicates possibly existing effect, *pd* > 97% indicates likely existing effect, *pd* > 99% indicates probably existing effect, and *pd* > 99.9% indicates almost certainly existing effect.

# Results

```{r}
#| label: data


```

## Sample description

```{r}
#| label: tbl-desc
#| tbl-cap: Demographic and descriptive clinical variables


```

## Cortical thickness

*4 Fíla*

## Subcortical volummetry


## Cognitive variables


# Appendix


# References

